\large How resource abundance and stochasticity affect organisms' range sizes

### 2025-02-17 writing/appendix-C-empirical-modeling.Rmd
rm "unaccounted"; reduced fig size

---
title: \large How resource abundance and stochasticity affect organisms' range sizes
subtitle: "Appendix C: Empirical modeling"
# author:
#   - name: Stefano Mezzini
#     email: stefano.mezzini@ubc.ca
#     institute: [biol, braes]
#   - name: Chris H. Fleming
#     institute: [ucf, smith]
#   - name: E. Patrícia Medici
#     institute: [tapir, iucn, escas]
#   - name: Michael J. Noonan
#     email: michael.noonan@ubc.ca
#     institute: [biol, braes, cmps]
# institute:
#   - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
#   - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
#   - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
#   - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
#   - tapir: Lowland Tapir Conservation Initiative (LTCI), Instituto de Pesquisas Ecológicas (IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
#   - iucn: IUCN SSC Tapir Specialist Group (TSG), Campo Grande, Brazil.
#   - escas: Escola Superior de Conservação Ambiental E Sustentabilidade (ESCAS/IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
#   - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'hr-resource-stoch.bib'
csl: 'the-american-naturalist.csl'
fontsize: 12pt
header-includes:
    - \renewcommand{\figurename}{Figure C\!\!} # for "Figure Sx.
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize} # smaller caption text
    -  \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    latex_engine: xelatex # for unicode characters
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
editor_options: 
  chunk_output_type: console
---

@@

```{r dag, eval=TRUE, echo=FALSE, fig.cap="Directed Acyclical Graph assumed for inferring the causal effects of $\\e(R)$ and $\\var(R)$ on $H$, where NDVI was used as a proxy for $R$. $Z$ and $S$ indicate unaccounted confounding factors that result from habitat-level variables (e.g., competition, predation, etc.) and satellite-level variables (e.g., noise, cloud cover).", fig.height=6, fig.width=7, out.width='\\linewidth'}
dagify(NDVI ~ S,
       EN ~ NDVI + E,
       VN ~ NDVI + V,
       E ~ R,
       V ~ R,
       Z ~ E + V,
       S ~ R,
       H ~ E + V + Z,
       outcome = 'H',
       labels = c(H = 'H', R = 'R', E = 'E(R)', V = 'Var(R)', S = 'S',
                  Z = 'Z', NDVI = 'NDVI', EN = 'E(NDVI)', VN = 'Var(NDVI)'),
       coords = list(x = c(R = 0, NDVI = 0, Z = 0, S = 0, H = 0,
                           E = -1, V = 1, EN = -1, VN = 1),
                     y = c(R = 3,
                           Z = 1.5, H = 0, E = 1.5, V = 1.5,
                           S = 4.5, NDVI = 6, EN = 4.5, VN = 4.5))) %>%
  tidy_dagitty() %>%
  mutate(col = case_when(name == 'E' ~ '1',
                         name == 'EN' ~ '1',
                         name == 'V' ~ '2',
                         name == 'VN' ~ '2',
                         name == 'H' ~ '3',
                         TRUE ~ '4')) %>%
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = col), size = 20, alpha = 0.3) +
  geom_dag_edges() +
  geom_dag_text(aes(label = label), color = 'black', size = 3) +
  scale_color_manual(values = c(pal[1:3], 'black')) +
  theme_dag() +
  theme(legend.position = 'none')
```

>>>>

---
title: \large How resource abundance and stochasticity affect organisms' range sizes
subtitle: "Appendix C: Empirical modeling"
# author:
#   - name: Stefano Mezzini
#     email: stefano.mezzini@ubc.ca
#     institute: [biol, braes]
#   - name: Chris H. Fleming
#     institute: [ucf, smith]
#   - name: E. Patrícia Medici
#     institute: [tapir, iucn, escas]
#   - name: Michael J. Noonan
#     email: michael.noonan@ubc.ca
#     institute: [biol, braes, cmps]
# institute:
#   - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
#   - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
#   - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
#   - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
#   - tapir: Lowland Tapir Conservation Initiative (LTCI), Instituto de Pesquisas Ecológicas (IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
#   - iucn: IUCN SSC Tapir Specialist Group (TSG), Campo Grande, Brazil.
#   - escas: Escola Superior de Conservação Ambiental E Sustentabilidade (ESCAS/IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
#   - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'hr-resource-stoch.bib'
fontsize: 12pt
header-includes:
    - \renewcommand{\figurename}{Figure C\!\!} # for "Figure Sx.
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize} # smaller caption text
    -  \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    latex_engine: xelatex # for unicode characters
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
editor_options: 
  chunk_output_type: console
---

@@

```{r dag, eval=TRUE, echo=FALSE, fig.cap="Directed Acyclical Graph assumed for inferring the causal effects of $\\e(R)$ and $\\var(R)$ on $H$, where NDVI was used as a proxy for $R$. $Z$ and $S$ indicate confounds that result from habitat-level variables (e.g., competition, predation, etc.) and satellite-level variables (e.g., noise, cloud cover).", fig.height=6, fig.width=6, out.width='75%'}
dagify(NDVI ~ S,
       EN ~ NDVI + E,
       VN ~ NDVI + V,
       E ~ R,
       V ~ R,
       Z ~ E + V,
       S ~ R,
       H ~ E + V + Z,
       outcome = 'H',
       labels = c(H = 'H', R = 'R', E = 'E(R)', V = 'Var(R)', S = 'S',
                  Z = 'Z', NDVI = 'NDVI', EN = 'E(NDVI)', VN = 'Var(NDVI)'),
       coords = list(x = c(R = 0, NDVI = 0, Z = 0, S = 0, H = 0,
                           E = -1, V = 1, EN = -1, VN = 1),
                     y = c(R = 3,
                           Z = 1.5, H = 0, E = 1.5, V = 1.5,
                           S = 4.5, NDVI = 6, EN = 4.5, VN = 4.5))) %>%
  tidy_dagitty() %>%
  mutate(col = case_when(name == 'E' ~ '1',
                         name == 'EN' ~ '1',
                         name == 'V' ~ '2',
                         name == 'VN' ~ '2',
                         name == 'H' ~ '3',
                         TRUE ~ '4')) %>%
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = col), size = 20, alpha = 0.3) +
  geom_dag_edges() +
  geom_dag_text(aes(label = label), color = 'black', size = 3) +
  scale_color_manual(values = c(pal[1:3], 'black')) +
  theme_dag() +
  theme(legend.position = 'none')
```
